---
title: "assignment"
author: "Yue Xiong"
date: '2022-07-27'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step1: Data Preparation

In step1, we aim to read in the original datasets and the synthetic dataset while aligning the variables especially set in the syn dataset. Attempting to compare the two datasets, we should also concatenate the original datasets vertically with variables all set as the same. Also, we are filtering out all the non-GPDR countries.

```{r echo=FALSE, include=FALSE}
# install the required R packages
# install.packages("readr")
# install.packages("vroom")
# install.packages("tidyverse")
# install.packages("arsenal")
# install.packages("reshape2")
# install.packages("synthpop")
library(readr)
library(vroom)
library(tidyverse)
library(arsenal)
library(reshape2)
library(synthpop)
```
```{r}
# set the working directory
setwd("F:/Master-Thesis-DifferentialPrivacy")
# read in the synthetic dataset
syn_data <- read.csv(file = "./syn_2020-08-02_2020-08-08.csv")
head(syn_data)
# syn_data <- read_csv("./syn_2020-08-02_2020-08-08.csv", show_col_types = FALSE)
# rename "sample weight" to "weight" to avoid comflicts
colnames(syn_data)[colnames(syn_data) == "sample_weight"] <- "weight"

# columns to be included from the original dataset
cols_list <- colnames(syn_data)

# initialize an empty dataset list
ori_dataset <- list()

for (i in 1:7){
ori_dataset[[i]] <- vroom(list.files(pattern = "*_full.csv$")[i],
                    show_col_types = FALSE) %>%
                    select(all_of(cols_list))
}

dim(ori_dataset[[2]])[2] == dim(syn_data)[2]
# check whether 2 dimensions coincide with each other

# bind the original datasets from 0802 to 0808 vertically
bindori_dataset <- bind_rows(ori_dataset)
bindori_dataset <- as.data.frame(bindori_dataset)
dim(bindori_dataset)
dim(syn_data)
```
Now we need to check all the `gpdr` countries and only do the alignment.
```{r}
gpdr_countries_data <- NA
gpdr_countries_data <- read.csv(file = "./gpdr.csv", sep = ",")
head(gpdr_countries_data$Country_GID, n = 10L)
```
We then filter the binded original datasets and syn dataset with Region_GID specified
```{r}
country_name <- unique(as.character(gpdr_countries_data$Country_GID))
length(country_name)
length(unique(bindori_dataset$GID_0))
bindori_dataset_filtered <- bindori_dataset %>%
                                        filter(GID_0 %in% country_name)


syn_data_filtered <- syn_data %>%
                          filter(GID_0 %in% country_name)
length(unique(bindori_dataset_filtered$GID_0)) == length(unique(syn_data_filtered$GID_0))  # check whether we have the same number of countries for both synthetic and original datasets.
```
## Step2: Evaluating the utility of the syn data
In step2, we try to evaluate the utility of the synthetic dataset with one-way marginal and two-way marginal measures. 

- As for the one-way utility, the syn dataset is measured with the compare plots and pMSE/S_pMSE.

- And for the two-way utility, the synthetic dataset is evaluated with the utility tables which takes up a heatmap fashion/manner.

```{r}
# subset some example columns and try plotting the compare histograms
# these are symptoms variables
symptoms <- c("B3","B4","B1_1","B1_2","B1_3","B1_4","B1_5","B1_6","B1_7",
              "B1_8","B1_9","B1_10","B1_12","B1_13",
              "B1b_x1","B1b_x2","B1b_x3","B1b_x4","B1b_x5","B1b_x6","B1b_x7",
              "B1b_x8","B1b_x9","B1b_x10","B1b_x12","B1b_x13","B2")

# these are testing variables
testing <- c("B7")

# Columns `B0`, `B8a`, `B15_1`, `B15_2`, `B15_3`, etc. don't exist

ori_dataset_symptoms <- as.data.frame(bindori_dataset_filtered[, symptoms])
syn_dataset_symptoms <- as.data.frame(syn_data_filtered[, symptoms])

ori_dataset_testing <- as.data.frame(bindori_dataset_filtered$B7)
syn_dataset_testing <- as.data.frame(syn_data_filtered$B7)
```

### (1). one-way marginals using compare()
```{r}
## for var B3 -> symptoms
compare(object = data.frame(B3 = syn_dataset_symptoms$B3),
        data = data.frame(B3 = ori_dataset_symptoms$B3),
        vars = c("B3"), cont.na = NULL,
        msel = NULL, stat = "percents", breaks = 20,
        nrow = 2, ncol = 2, rel.size.x = 1,
        utility.stats = c("pMSE", "S_pMSE"),
        cols = c("#1A3C5A","#4187BF"),
        plot = TRUE, table = TRUE)
```
```{r}
## for var B1_1 -> symptoms
compare(object = data.frame(B1_1 = syn_dataset_symptoms$B1_1),
        data = data.frame(B1_1 = ori_dataset_symptoms$B1_1),
        vars = "B1_1", cont.na = NULL,
        msel = NULL, stat = "percents", breaks = 20,
        nrow = 2, ncol = 2, rel.size.x = 1,
        utility.stats = c("pMSE", "S_pMSE"),
        cols = c("#1A3C5A","#4187BF"),
        plot = TRUE, table = FALSE)
```

```{r}
## for var B7 -> testing
compare(object = data.frame(B7 = syn_data_filtered$B7),
        data = data.frame(B7 = bindori_dataset_filtered$B7),
        vars = "B7", cont.na = NULL,
        msel = NULL, stat = "percents", breaks = 20,
        nrow = 2, ncol = 2, rel.size.x = 1,
        utility.stats = c("pMSE", "S_pMSE"),
        cols = c("#1A3C5A","#4187BF"),
        plot = TRUE, table = FALSE)
```

```{r}
## for var D1 -> module B
compare(object = data.frame(D1 = syn_data_filtered$D1),
        data = data.frame(D1 = bindori_dataset_filtered$D1),
        vars = "D1", cont.na = NULL,
        msel = NULL, stat = "percents", breaks = 20,
        nrow = 2, ncol = 2, rel.size.x = 1,
        utility.stats = c("pMSE", "S_pMSE"),
        cols = c("#1A3C5A","#4187BF"),
        plot = TRUE, table = FALSE)
```

```{r}
## for var D2 -> module B
compare(object = data.frame(D2 = syn_data_filtered$D2),
        data = data.frame(D2 = bindori_dataset_filtered$D2),
        vars = "D2", cont.na = NULL,
        msel = NULL, stat = "percents", breaks = 20,
        nrow = 2, ncol = 2, rel.size.x = 1,
        utility.stats = c("pMSE", "S_pMSE"),
        cols = c("#1A3C5A","#4187BF"),
        plot = TRUE, table = FALSE)
```
```{r}
## for var E2 -> demographics
compare(object = data.frame(E2 = syn_data_filtered$E2),
        data = data.frame(E2 = bindori_dataset_filtered$E2),
        vars = "E2", cont.na = NULL,
        msel = NULL, stat = "percents", breaks = 20,
        nrow = 2, ncol = 2, rel.size.x = 1,
        utility.stats = c("pMSE", "S_pMSE"),
        cols = c("#1A3C5A","#4187BF"),
        plot = TRUE, table = FALSE)
```

```{r}
## for var E3 -> demographics
compare(object = data.frame(E3 = syn_data_filtered$E3),
        data = data.frame(E3 = bindori_dataset_filtered$E3),
        vars = "E3", cont.na = NULL,
        msel = NULL, stat = "percents", breaks = 20,
        nrow = 2, ncol = 2, rel.size.x = 1,
        utility.stats = c("pMSE", "S_pMSE"),
        cols = c("#1A3C5A","#4187BF"),
        plot = TRUE, table = FALSE)
```

